\documentclass[bsc, frontabs, twoside, singlespacing, parskip]{infthesis}

\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{url}
%\usepackage[font=small,format=plain,labelfont=bf,up,textfont=it,up]{caption}
%\usepackage{fullpage}
%\usepackage{parskip}
\usepackage{listings}
\usepackage{color}
\usepackage{subcaption}
\usepackage{acronym}
\usepackage{todonotes}
\usepackage{afterpage}
\usepackage{pdflscape}
\usepackage{morefloats}
\usepackage{pdfpages}


\let\oldtodo\todo
\newcommand{\tdi}[1]{\oldtodo[inline]{#1}}
\renewcommand\todo[1]{\oldtodo[size=\tiny]{#1}}
\newcommand{\tdic}[2][]{
    \oldtodo[inline, caption={#1}]{
        \begin{minipage}{
            \textwidth-4pt
        }
        #2
        \end{minipage}
    }
}
\setlength{\marginparwidth}{2cm}

\definecolor{mygreen}{rgb}{0,0.6,0}
\definecolor{mygray}{rgb}{0.5,0.5,0.5}
\definecolor{mymauve}{rgb}{0.58,0,0.82}

\lstset{ %
  backgroundcolor=\color{white},   % choose the background color; you must add \usepackage{color} or \usepackage{xcolor}
  basicstyle=\footnotesize,        % the size of the fonts that are used for the code
  breakatwhitespace=false,         % sets if automatic breaks should only happen at whitespace
  breaklines=true,                 % sets automatic line breaking
  captionpos=b,                    % sets the caption-position to bottom
  commentstyle=\color{mygreen},    % comment style
  deletekeywords={...},            % if you want to delete keywords from the given language
  escapeinside={\%*}{*)},          % if you want to add LaTeX within your code
  extendedchars=true,              % lets you use non-ASCII characters; for 8-bits encodings only, does not work with UTF-8
  keywordstyle=\color{blue},       % keyword style
  language=Python,                 % the language of the code
  morekeywords={*,...},            % if you want to add more keywords to the set
  numbers=left,                    % where to put the line-numbers; possible values are (none, left, right)
  numbersep=5pt,                   % how far the line-numbers are from the code
  numberstyle=\tiny\color{mygray}, % the style that is used for the line-numbers
  rulecolor=\color{black},         % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. comments (green here))
  showspaces=false,                % show spaces everywhere adding particular underscores; it overrides 'showstringspaces'
  showstringspaces=false,          % underline spaces within strings only
  showtabs=false,                  % show tabs within strings adding particular underscores
  stepnumber=1,                    % the step between two line-numbers. If it's 1, each line will be numbered
  stringstyle=\color{mymauve},     % string literal style
  tabsize=2,                       % sets default tabsize to 2 spaces
}

\acrodef{UI}{User Interface}
\acrodef{UX}{User Experience}
\acrodef{tf.idf}{term frequency-inverse document frequency}
\acrodef{GUI}{Graphical User Interface}
\acrodef{API}{Application Programming Interface}
\acrodef{MVC}{Model View Controller}
\acrodef{HCI}{Human Computer Interaction}
\acrodef{RPC}{Remote Procedure Call}

\begin{document}

\title{A Tool for Visualising Cell Model Results \\
       MInf Project Phase 2}
\author{James Hulme}
\course{Master of Informatics}
\project{MInf Project (Phase 2) Report}
\abstract{
This project was the creation of a tool to help biologists visualise and analyse their data better.  The project was developed in two phases.  This document details the second phase.  In this phase of the project: animation of cell data was added, along with functionality for annotating visualisations, normalising and exporting data, using plots as queries, and real time collaboration.  The project was evaluated using a mixture of user feedback, relevant metrics, and experimentation.  The findings indicated that the tool that was developed was an effective tool which fills a gap left by the currently available software.
}

\maketitle

\section*{Acknowledgements}
I would like to thank my supervisor, Professor Jane Hillston, for all the support and advice throughout the past two years.  Particular thanks is due for my improved writing ability. Early drafts of my writing were difficult to parse, but she made it past that and has greatly improved my writing ability.

I would also like to thank Dr Vashti Galpin for all the Bio-PEPA support she has provided over the course of this project.  The projet would not have been possible without here willingness to provide a range of models.  Thanks is also due for being a repeated evaluation user.  Her feedback on the tool has been invaluable.

Finally, I would like to thank Dr Lorraine Kerr and Dr Zeenat Noordally for repeatedly giving up their valuable time and being evaluation participants.  The project would not be as useful as it now is without their most helpful feedback.

\thispagestyle{empty}

\newpage

\thispagestyle{empty}

\newpage

\tableofcontents
\thispagestyle{empty}

\clearpage
\setcounter{page}{1}

\input{intro}
\input{goals}
\input{work}
\input{evaluation}
%\input{workplan}
\input{conclusion}

%\clearpage
%\addcontentsline{toc}{section}{References}
\bibliographystyle{unsrt}
\bibliography{biblio}

\clearpage
\chapter{Appendix}
\input{appendix}

\end{document}
